Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2013-11-13 paper/paper.tex
rewrote equation and other things in Model and Methods, moved some of what was there to the pill section Tested everything on node23.

\section{Model and Methods}
We implement the model of Huang \emph{et al.}  for the behavior of the
MinD and MinE proteins inside the cell~\cite{huang2003dynamic}.  This
model is defined by a set of five reaction-diffusion equations:
\begin{align}
  \frac{\partial n_{?}}{\partial t} &= \text{\davidsays{Actual differential equations here.}}
\end{align}
These equations define the rates of diffusion and reaction shown in
Figure~\ref{fig:reactions}.  \davidsays{Add more text here, describing
  the system and the reactants.  In particular, we want to state that
  MinD:ADP binds ATP (or is hydrolyzed?... check!) to form MinD:ATP,
  which then gets stuck to the cell membrane, which is more likely
  when there is already MinD:ATP bound to the membrane.  The MinE then
  binds to the membrane-bound MinD:ATP.  Finally the whole complex can
  fall off the membrane, at which stage you end up with MinD:ADP and
  are ready to start over.  This is also where we define our
  terminology for the volume and surface densities.}

@@

We start our cells with the MinD and MinE concentrations that are
reported as wild type concentrations by Huang et al . We as well use the same reaction
diffusion constants and and reaction rates,
\begin{gather*} %format better
  \mathcal{D}_D = \mathcal{D}_{E}  = 2.5 \micron^2/\text{sec}, \\
  \sigma_D^{\textrm{ADP $\rightarrow$ ATP}}  = 1/\textrm{sec},  \sigma_D = 0.025 \micron /\textrm{sec}, \\
  \sigma_{dD}  = 0.0015 \micron^3/ \textrm{sec}, \\
  \sigma_{de}  = 0.7/\textrm{sec}, \sigma_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}

>>>>

\section{Model and Methods}
We implement the model of Huang \emph{et al.}  for the behavior of the
MinD and MinE proteins inside the cell~\cite{huang2003dynamic}.
Figure~\ref{fig:reactions} shows the reaction process.  The
cytoplasmic MinD:ADP complex undergoes nucleotide exchange and is
changed into the MinD:ATP complex.  This will naturally diffuse
towards and attach to the cell membrane.  A cytoplasmic MinE will
attach to the wall bound MinD:ATP complex and after a time will
activate ATP hydrolosis.  This breaks up the complex, releasing MinE,
phosphate, and MinD:ADP back into the cell.  The MinD:ADP will undergo
nucleotide exchange and begin again the cyclic process.  There is a
mechanism for MinD:ATP clustering that is shown in
Equation~\ref{fig:FourthPDE}.  When brushing up against the cell
membrane, the MinD:ATP attachement rate increases linearly with the
density of MinD:ATP already attached.  This model is defined by a set of
five reaction-diffusion equations:

\begin{multline}
  \frac{\partial \rho_{D:ADP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ADP}-k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  +\delta(d_w)k_{de}\sigma_{DE}\hspace{3.4cm}
\end{multline}
\begin{multline}
  \frac{\partial \rho_{D:ATP}}{\partial t} = \mathcal{D}_D\nabla^2\rho_{D:ATP}+k_D^{ADP\rightarrow ATP}\rho_{D:ADP}\\
  -\delta(d_w)[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
\end{multline}
\begin{multline}
  \frac{\partial \rho_E}{\partial t} = \mathcal{D}_E\nabla^2\rho_E+\delta(d_w)k_{de}\sigma_{DE}
  -\delta(d_w)k_E \sigma_D \rho_E
\end{multline}
\begin{multline}
  \frac{\partial \sigma_d}{\partial t} = -k_E\rho_d\rho_E
  +[k_D+k_{dD}(\sigma_D+\sigma_{DE})]\rho_{D:ATP}
  \label{fig:FourthPDE}
\end{multline}
\begin{multline}
  \frac{\partial \sigma_{DE}}{\partial t} = -k_{de}\sigma_{DE}+k_E\sigma_D\rho_E\hspace{3cm}
  \label{fig:FifthPDE}
\end{multline}

Where $\rho$ is cytoplasmic protein density ($\mu m^{-3}$), $\sigma$
is membrane bound density ($\mu m^{-2}$), $\mathcal{D}_D$ and
$\mathcal{D}_{E}$ th diffusions constants for MinD and MinE,
respectively, $k_D^{\textrm{ADP $\rightarrow$ ATP}}$ the rate of
conversion from MinD:ADP to the MinD:ATP complex, $k_D$ the rate of
MinD:ATP attachement to the membrane when no protein is already
attached there, $k_{dD}$ the increase of this rate when MinD:ATP is
present on the membrane, $k_{de}$ the rate of hydrolisis of the
MinD:MinE:ATP complex, $k_E$ the rate of cytoplasmic MinE binding to
membrane bound MinD:ATP complex, and $d_w$ is the distance from the
point in space to the closest wall (which will always be
perpendicular to this distance).  $\delta(d_w)$ has units of
$\mu m^{-1}$ and will be zero everywhere except at the wall.
Equations \ref{fig:FourthPDE} and \ref{fig:FifthPDE} will only be
relevent at the membrane because the membrane bound density values
will have no meaning in the cytoplasm where there is no membrane.

Our diffusion and reaction rates are shown below.  We are interested
primarily with the effect of cellular shape on the protein
oscillations, so we follow Huang's work\cite{huang2003dynamic} and do
not deviate from these values.  Huang's simulations use total MinD and
MinE concentrations of $1,000/\mu m$ and $350/\mu m$, respectively, in
a cylindrical cell of radius $0.5\mu m$, and in our (non-cylindrical)
cells we use the same number of proteins per unit volume.

\begin{gather*}
  \mathcal{D}_D = \mathcal{D}_{E} = 2.5\micron^2/\text{sec}\\
  k_D^{\textrm{ADP $\rightarrow$ ATP}} = 1/\textrm{sec,  }
  k_D = 0.025 \micron /\textrm{sec}\\
  k_{dD} = 0.0015 \micron^3/ \textrm{sec,  }
  k_{de} = 0.7/\textrm{sec}\\
  k_E = 0.093 \micron^3 /\textrm{sec}.
\end{gather*}

@@

We start our cells with the MinD and MinE concentrations that are
reported as wild type concentrations by Huang et al . We as well use the same reaction
diffusion constants and and reaction rates,
